Choosing Color Palettes for Statistical Graphics

### 2006-05-29 trunk/vcd/paper/Color/draft.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/vcd@540 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<votes, fig=TRUE, height=9, width=11>>=
votes <- votes[c(1, 3, 4, 5, 9, 11, 13, 14, 15, 16, 2, 6, 7, 8, 10, 12),]
votes <- votes[,c(2, 4, 1, 3, 5)]
votes <- as.table(as.matrix(votes))
names(dimnames(votes)) <- c("", "")

@@

To illustrate qualitative palettes, data from the 2005 election for the German parliament `Bundestag'
are employed. In that election, five parties were able to obtain enough votes to enter the
Bundestag---the distribution of seats is depicted in a pie chart in Figure~\ref{fig:seats}.
The colors used are rough metaphors for the political parties, using a red hue $H = 0$ for the
social democrats SPD, a blue hue $H = 240$ for the conservative CDU, a yellow hue $H = 60$
for the liberal FDP, a green hue $H = 120$ for the green party `Die Gruenen' and a purple 
hue $H = 300$ for the leftist party `Die Linke'. All colors use the same chroma $C = 60$ and
luminance $L = 75$. The pie chart clearly shows that neither the governing coalition of SPD
and Gruene nor the opposition of CDU and FDP could assemble a majority. Given that no party
would enter a coalition with the leftists, this lead to a big coalition of CDU and SPD.
Figure~\ref{fig:votes} shows the distribution of votes in this election stratified by 
province (Bundesland) in a highlighted mosaic display. The order of provinces is from 
north to south, first for the 10 Western provinces, then for the 6 Eastern provinces. Clearly,
the SPD performed better in the North and the CDU better in the South; furthermore, the
Die Linke performed particularly well in the Eastern provinces and the Saarland.

>>>>

\setkeys{Gin}{width=.8\textwidth}
\begin{figure}[p]
\begin{center}
<<votes, fig=TRUE, height=9, width=11>>=
votes <- votes[c(1, 3, 4, 5, 9, 11, 13, 14, 15, 16, 2, 6, 7, 8, 10, 12),]
votes <- votes[,c("CDU/CSU", "FDP", "SPD", "Gruene", "Linke")]
votes <- as.table(as.matrix(votes))
names(dimnames(votes)) <- c("", "")

@@

To illustrate qualitative palettes, data from the 2005 election for the German parliament `Bundestag'
are employed. In that election, five parties were able to obtain enough votes to enter the
Bundestag---the distribution of seats is depicted in a pie chart in Figure~\ref{fig:seats}.
The colors used are rough metaphors for the political parties, using a red hue $H = 0$ for the
social democrats SPD, a blue hue $H = 240$ for the conservative CDU/CSU, a yellow hue $H = 60$
for the liberal FDP, a green hue $H = 120$ for the green party `Die Gruenen' and a purple 
hue $H = 300$ for the leftist party `Die Linke'. All colors use the same chroma $C = 60$ and
luminance $L = 75$. The pie chart clearly shows that neither the governing coalition of SPD
and Gruene nor the opposition of CDU and FDP could assemble a majority. Given that no party
would enter a coalition with the leftists, this lead to a big coalition of CDU and SPD.
Figure~\ref{fig:votes} shows the distribution of votes in this election stratified by 
province (Bundesland) in a highlighted mosaic display. The order of provinces is from 
north to south, first for the 10 Western provinces, then for the 6 Eastern provinces. Clearly,
the SPD performed better in the North and the CDU better in the South; furthermore, the
Die Linke performed particularly well in the Eastern provinces and the Saarland.
